---
title: "EDLD 650 Review Session"
author: "David D. Liebowitz"
output:
  xaringan::moon_reader:
  css: ['default', 'rladies', 'ki-fonts', 'my_custom.css', 'xaringanthemer.css']
# self_contained: true
nature:
  highlightStyle: github
highlightLines: true
countIncrementalSlides: false
editor_options: 
  
  chunk_output_type: console
---
```{css, echo=F}
.inverse {
  background-color : #2293bf;
}
```

```{R, setup, include = F}
library(pacman)

p_load(MASS, tidyverse, fixest, lme4, ggplot2, xaringan, knitr, xaringanthemer, ggthemes)

# Define color
extra_css <- list(
  ".red"   = list(color = "red"),
  ".blue"  =list(color = "blue"),
  ".green" = list(color = "#8bb174"),
  ".purple" = list(color = "#6A5ACD"),
  ".red-pink" = list(color= "#e64173"),
  ".grey-light" = list(color= "grey70"),
  ".slate" = list(color="#314f4f"),
  ".small" = list("font-size" = "90%"),
  ".large" = list("font-size" = "120%"))
write_extra_css(css = extra_css, outfile = "my_custom.css")
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 6.75,
  fig.width = 10.5,
  warning = F,
  message = F
)
# opts_chunk$set(dev = "svg")
# 
# options(device = function(file, width, height) {
#   svg(tempfile(), width = width, height = height)
# })
options(knitr.table.format = "html")
hook_output <- knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
  lines <- options$output.lines
  if (is.null(lines)) {
    return(hook_output(x, options))  # pass to default hook
  }
  x <- unlist(strsplit(x, "\n"))
  more <- "..."
  if (length(lines)==1) {        # first n lines
    if (length(x) > lines) {
      # truncate the output, but add ....
      x <- c(head(x, lines), more)
    }
  } else {
    x <- c(more, x[lines], more)
  }
  # paste these lines together
  x <- paste(c(x, ""), collapse = "\n")
  hook_output(x, options)
})




```


# SAT score and income

```{r, echo=F, fig.height=5.5}
set.seed(5)

### build School1

mu <- c(475,475)

sigma <- rbind(c(100,0.3),c(0.3,100))

g1 <- as.data.frame(mvrnorm(n=1000, mu=mu, Sigma=sigma))

g1$school <- c("School 1")


### build School 2

mu <- c(500,500)

sigma <- rbind(c(100,-0.7),c(-0.7,100) )

g2<-as.data.frame(mvrnorm(n=1000, mu=mu, Sigma=sigma))

g2$school <- c("School 2")



### build School 3

mu<-c(525,525)

sigma<-rbind(c(150,-0.7),c(-0.7,150) )

g3<-as.data.frame(mvrnorm(n=1000, mu=mu, Sigma=sigma))

g3$school <- c("School 3")


# the combined data of all three groups
df<-rbind(g1,g2,g3)

df <- df %>% mutate(income30 = 100*V2)
df <- rename(df, SAT = V1)
df$school <- as_factor(df$school)

df %>% 
  ggplot(aes(x=SAT, y=income30))+geom_point()+ geom_smooth(method='lm') +
  theme_pander(base_size=18)
```
---
# A closer look

```{r, echo=F, fig.height=5.5}
df %>%
  ggplot(aes(x=SAT, y=income30, group=school, col=school))+geom_point() +
  theme_pander(base_size=18)
```
---
# A closer look
```{r, echo=F, fig.height=5.5}
df %>%
  ggplot(aes(x=SAT, y=income30, group=school, col=school))+geom_point() + geom_smooth(method='lm', col='black') +
  theme_pander(base_size=18)
```

---
# Graphical fixed effects

```{r, echo=T}
df <- df %>% group_by(school) %>% mutate(mean_SAT = mean(SAT))
df <- df %>% mutate(demean_SAT = SAT - mean_SAT)
```

--

```{r, echo=F, fig.height=5.5}

df %>%
  ggplot(aes(x=demean_SAT, y=income30, group=school, col=school))+geom_point() + geom_smooth(method='lm', col='black') +
  theme_pander(base_size=18)
```

---
# Graphical fixed effects

```{r, echo=T}
df <- df %>% group_by(school) %>% mutate(mean_income=mean(income30))
df <- df %>% mutate(demean_income = income30 - mean_income)
```

--

```{r, echo=F, fig.height=5.5}
df %>%
  ggplot(aes(x=demean_SAT, y=demean_income, group=school, col=school))+geom_point() + 
  theme_pander(base_size=18)

```

---
# Graphical fixed effects

```{r, echo=T}
df <- df %>% group_by(school) %>% mutate(mean_income=mean(income30))
df <- df %>% mutate(demean_income = income30 - mean_income)
```

```{r, echo=F, fig.height=5.5}
df %>%
  ggplot(aes(x=demean_SAT, y=demean_income, group=school, col=school))+geom_point() + 
  geom_smooth(aes(x=demean_SAT, y=demean_income), method='lm', col='black', inherit.aes = F) +
  theme_pander(base_size=18)

```

---
# A naive regression
```{r, echo=T, output.lines=c(1:14)}
summary(lm(income30 ~ SAT, data=df))
```

--

$$INCOME_{i} = \beta_{0} + \beta_{1} SAT_{i} + \varepsilon_{i}$$

---
# Adjust for school
```{r, echo=T, output.lines=c(3:16)}
summary(lm(income30 ~ SAT + school, data=df))
```

--

$$INCOME_{ij} = \beta_{0} + \beta_{1} SAT_{ij} + \beta_{2} SCHOOL2_{j} + \beta_{3} SCHOOL3_{j} + \varepsilon_{i}$$
---
# Cluster-adjusted SEs
```{r, echo=T}
summary(fixest::feols(income30 ~ SAT | school, data=df))
```

--

**What happened to the standard errors? Is that right? Why or why not?**

---
# Fixed vs. random

### Fixed:
$$INCOME_{ij} = \beta_{0} + \beta_{1} SAT_{ij} + \beta_{2} SCHOOL2_{j} + \beta_{3} SCHOOL3_{j} + \varepsilon_{ij}$$

--

### Random:
$$INCOME_{ij} = \gamma_{0} + \gamma_{1} SAT_{ij} + (\nu_{j} + \varepsilon_{ij})$$

---
### Random intercepts
$$INCOME_{ij} = \gamma_{0} + \gamma_{1} SAT_{ij} + \color{red}{(\nu_{j} + \varepsilon_{ij})}$$

There is now a composite residual that has parts *for each individual*. A school-level residual $(\nu_{j})$ and an individual-level residual $(\varepsilon_{ij})$. Each residual term can be summarized by examining its variance/SD and correlated with other residual terms.

--

$\nu_{j}$: unique to school *j*; identical for all students in *jth* school; iid for every school

$\varepsilon_{ij}$: unique to student i in school j. iid for every student

--

Can calculate the **intra-class correlation**, *a summary of the proportion of outcome variability attributable to differences across schools.*

$$\hat{\rho} = \frac{\sigma^{2}_{s}}{\sigma^{2}_{s}+\sigma^{2}_{i}}$$

--

.small[Random effects estimation procedures use iterative algorithms (Maximum Likelihood Estimation, restricted MLE, Gradient descent, etc.) to minimize the sum of the square errors at each hierarchical level of the model.]

---
# Estimate random intercept model
```{r, echo=T}
summary(lme4::lmer(income30 ~ SAT + (1 | school), data=df))
```

---
# Treatment effect?
```{r, echo=F, output.lines=-c(1:6)}
summary(lme4::lmer(income30 ~ SAT + (1 | school), data=df))
```

--

**What's different/same compared to fixed effect estimates?**
